---
title: "Reverse Engineering Project"
author: "Danny Martella"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
majordata <- read_csv("https://umd.instructure.com/files/88374270/download?download_frd=1&verifier=OZV0ivOyNlWt6XKhlwZ3x6Z1NjNkhaDdt0AOhCD1")
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

  clean_names(majordata)
  
 majordata_dollar <- majordata |>
    mutate(
     income_dollar = case_when(
        str_detect(income, "1") ~ 9999,
        str_detect(income, "2") ~ 19999,
        str_detect(income, "3") ~ 29999,
        str_detect(income, "4") ~ 39999,
        str_detect(income, "5") ~ 49999,
        str_detect(income, "6") ~ 59999,
        str_detect(income, "7") ~ 69999,
        str_detect(income, "8") ~ 79999,
        str_detect(income, "9") ~ 89999,
        str_detect(income, "10") ~ 99999,
        str_detect(income, "11") ~ 109999,
        str_detect(income, "12") ~ 119999,
        str_detect(income, "13") ~ 129999,
        str_detect(income, "14") ~ 139999,
        str_detect(income, "15") ~ 149999,
        str_detect(income, "16") ~ 159999,
        .default = NA)
    )

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

# **Sentence text**: Hindus report the highest income of any tradition—around $85,000 per year—while three branches of Judaism (Reform, Conservative, and “Other Jew”) all average above $80,000. Episcopalians and members of the PCUSA also land in this high-income, high-education quadrant, both averaging above $75,000.
# **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]



```{r}
# Put code to reverse engineer sentence here



majordata|>
  group_by(trad) |>
  summarise(
    avg_income = round(mean(income)) 
  ) |>
  arrange(desc(avg_income))

majordata_dollar|>
   group_by(trad) |>
  summarise(
    avg_income_dollar = round(mean(income_dollar)) 
  ) |>
  arrange(desc(avg_income_dollar))
# Display results of code below this codeblock

```
# When I do not add dollar amounts to the income values, the rankings come out similar to the sentence from the article, with similar values in the 16 point scale. However, when I try to add dollar amounts before making the rankings, my list comes out in almost reverse order. 

### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-