---
title: "Reverse Engineering Project"
author: "Danny Martella"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, Why Religion, Not Income, Predicts the American Vote https://thescoop.org/religion/reverse-engineering/predicting-vote.html 

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
majordata <- read_csv("https://umd.instructure.com/files/88374270/download?download_frd=1&verifier=OZV0ivOyNlWt6XKhlwZ3x6Z1NjNkhaDdt0AOhCD1")
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

  clean_names(majordata)
  
 majordata_dollar <- majordata |>
    mutate(
     income_dollar = case_when(
        str_detect(income, "1") ~ 9999,
        str_detect(income, "2") ~ 19999,
        str_detect(income, "3") ~ 29999,
        str_detect(income, "4") ~ 39999,
        str_detect(income, "5") ~ 49999,
        str_detect(income, "6") ~ 59999,
        str_detect(income, "7") ~ 69999,
        str_detect(income, "8") ~ 79999,
        str_detect(income, "9") ~ 89999,
        str_detect(income, "10") ~ 99999,
        str_detect(income, "11") ~ 109999,
        str_detect(income, "12") ~ 119999,
        str_detect(income, "13") ~ 129999,
        str_detect(income, "14") ~ 139999,
        str_detect(income, "15") ~ 149999,
        str_detect(income, "16") ~ 159999,
        .default = NA)
    )

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

# **Sentence text**: Hindus report the highest income of any tradition—around $85,000 per year—while three branches of Judaism (Reform, Conservative, and “Other Jew”) all average above $80,000. Episcopalians and members of the PCUSA also land in this high-income, high-education quadrant, both averaging above $75,000.
# **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]



```{r}
# Put code to reverse engineer sentence here



majordata|>
  group_by(trad) |>
  summarise(
    avg_income = round(mean(income)) 
  ) |>
  arrange(desc(avg_income))

majordata_dollar|>
   group_by(trad) |>
  summarise(
    avg_income_dollar = round(mean(income_dollar)) 
  ) |>
  arrange(desc(avg_income_dollar))
# Display results of code below this codeblock

```
# When I do not add dollar amounts to the income values, the rankings come out similar to the sentence from the article, with similar values in the 16 point scale. However, when I try to add dollar amounts before making the rankings, my list comes out in almost reverse order. 

### Sentence 2

* **Sentence text**: "Trump pulled support from over 80% of white evangelicals and under 10% of Black Protestants—even when comparing respondents with the same household income."
* **Analysis summary**: Based on the code below, I think I got pretty close to what the article was talking about in this sentence. My WEpctT is the percentage of White Evangelicals that voted for Trump and BPpctT is the percentage of Black Protestants that voted for Trump. As you can see we get a figure at around 80% for White Evangelicals and a little over 10% for Black Protestants so I think I'm in the right ballpark. In order to get this close, I had to make my total for each religion those who voted for either Trump or Harris instead of the true total. Then we prove the income part of that statement below when we get very similar percentages when we separate to lower income (LI) and upper income (UI). 

```{r}
# Put code to reverse engineer sentence here
WE <- majordata |> filter(trad == "White Evangelical" & vote_2024 < 3)
WE_T <- majordata |> filter(trad == "White Evangelical" & vote_2024 == 2)
BP <- majordata |> filter(trad == "Black Protestant" & vote_2024 < 3)
BP_T <- majordata |> filter(trad == "Black Protestant" & vote_2024 == 2)

  WEpctT = (nrow(WE_T) / nrow(WE))
  BPpctT = (nrow(BP_T) / nrow(BP))


# Display results of code below this codeblock
WEpctT
BPpctT
  
```


```{r}
WE_LI <- majordata |> filter(trad == "White Evangelical" & vote_2024 < 3 & income < 9)
WE_T_LI <- majordata |> filter(trad == "White Evangelical" & vote_2024 == 2 & income < 9)
BP_LI <- majordata |> filter(trad == "Black Protestant" & vote_2024 < 3 & income < 9)
BP_T_LI <- majordata |> filter(trad == "Black Protestant" & vote_2024 == 2 & income < 9)

  WEpctT_LI = (nrow(WE_T_LI) / nrow(WE_LI))
  BPpctT_LI = (nrow(BP_T_LI) / nrow(BP_LI))
  
  
  WE_UI <- majordata |> filter(trad == "White Evangelical" & vote_2024 < 3 & income > 8)
WE_T_UI <- majordata |> filter(trad == "White Evangelical" & vote_2024 == 2 & income > 8)
BP_UI <- majordata |> filter(trad == "Black Protestant" & vote_2024 < 3 & income > 8)
BP_T_UI <- majordata |> filter(trad == "Black Protestant" & vote_2024 == 2 & income > 8)

  WEpctT_UI = (nrow(WE_T_UI) / nrow(WE_UI))
  BPpctT_UI = (nrow(BP_T_UI) / nrow(BP_UI))
```

### Sentence 3

* **Sentence text**: "Among non-white evangelicals who make at least $500K per year, they are only 1.2 percentage points more likely to vote for Donald Trump than members of the same group making less than $10K per year. And that’s actually the largest gap I found."
* **Analysis summary**: When I do a similar sequence to the previous sentence, because the sentences are similar, I do not get the 1.2 percent gap that he does. I am getting 100% for the Nonwhite evangelicals who make at least 500k and voted for trump, likely because there is only two of them. 

```{r}
# Put code to reverse engineer sentence here
NWE_500 <- majordata |> filter(trad == "Nonwhite Evangelical" & vote_2024 < 3 & income == 16)
NWE_T_500 <- NWE_500 |> filter(vote_2024 == 2)

NWE_10 <- majordata |> filter(trad == "Nonwhite Evangelical" & vote_2024 < 3 & income == 1)
NWE_T_10 <- NWE_10 |> filter(vote_2024 == 2)
# Display results of code below this codeblock
 NWEpctT_500 = (nrow(NWE_T_500) / nrow(NWE_500))
  NWEpctT_10 = (nrow(NWE_T_10) / nrow(NWE_10))
```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
