---
title: "lab_06"
author: "Danny Martella"
date: "2026-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# load your libraries
library(tidyverse)
library(lubridate)
```

## Getting Started

This week we'll be combining and joining data to make it more useful and to ask some questions about it. We'll use clergy contribution data from multiple election cycles and United Methodist church membership data to do this.

## Load Data

First, let's load all the data we'll need for this lab.

```{r}
# Load the five clergy contribution files for different election cycles
clergy_contribs_2016 <- read_csv("data/clergy_contribs_2016.csv")
clergy_contribs_2018 <- read_csv("data/clergy_contribs_2018.csv")
clergy_contribs_2020 <- read_csv("data/clergy_contribs_2020.csv")
clergy_contribs_2022 <- read_csv("data/clergy_contribs_2022.csv")
clergy_contribs_2024 <- read_csv("data/clergy_contribs_2024.csv")

# Load the two United Methodist data files
umdata_people <- read_csv('data/umdata_people.csv')
umdata_people_all <- read_csv('data/umdata_people_all.csv')
```

## Answer questions

Q1. Now you're interested in understanding giving patterns by election cycle. Combine the five clergy contribution files into a single dataframe, then calculate the total dollar amount contributed in each election cycle, and the total number of contributions in each cycle. Which cycle had the most money contributed? Which had the most individual contributions? What's the difference between those two results, and what might explain it?

A1. 

```{r}
clergy_contribs_combined <- bind_rows(list(clergy_contribs_2016, clergy_contribs_2018, clergy_contribs_2020, clergy_contribs_2022, clergy_contribs_2024))
View(clergy_contribs_combined)
clergy_totals_by_cycle <- clergy_contribs_combined |>
  group_by(contributor_name, two_year_transaction_period) |>
  summarize(
    total_amount = sum(contribution_receipt_amount),
    num_contributions = n(),
  )
clergy_wide <- clergy_totals_by_cycle |>
  select(contributor_name, two_year_transaction_period, total_amount) |>
  pivot_wider(
    names_from = two_year_transaction_period,
    values_from = total_amount,
    values_fill = 0
  )

```
```{r}
clergy_totals_by_cycle |>
  group_by(two_year_transaction_period) |>
  summarise(totalmoney = sum(total_amount), totalcont = sum(num_contributions))
```
#2024 had the most amount of contributions and most amount of money contributed. This could be because the same people that have been donating every cycle are getting older and maybe less aware of the money they are donating, or more people were donating this cycle. 

Q2. Some clergy members have been more consistent donors across election cycles. Using the combined contribution data, create a dataframe that shows each contributor and how many different election cycles they contributed in (someone who gave in 2016, 2018, and 2024 would have 3 cycles). You can use group_by and summarize to do this. Who contributed in all five election cycles? How many people contributed in only one cycle vs. multiple cycles?

A2. 

```{r}
num_years <- clergy_wide |>
 group_by(contributor_name) |>
   summarise( years_cont = sum(`2024` > 0, `2022` > 0, `2020` > 0, `2018` > 0,`2016` > 0)) |>
  arrange(desc(years_cont)) 
  
num_years |>
 filter(years_cont > 1) 

num_years |>
  filter(years_cont == 1)
   
```
# 665 people contributed in more than 1 election cycle, 1417 people contributed in just one cycle. 


Q3. You're interested in whether there are geographic patterns in clergy giving. Using the combined contribution data, calculate the total amount contributed by clergy in each city (`contributor_city`). What are the top 5 cities by total contribution amount? Then calculate the number of unique contributors in each of those top 5 cities. Are the cities with the most money also the cities with the most contributors, or are there differences?

A3. 

```{r}
clergy_contribs_combined |>
  group_by(contributor_city) |>
  summarise(city_sum = sum(contribution_receipt_amount))|>
arrange(desc(city_sum))
```

Q4. Use `pivot_wider` to create a summary showing each contributor and their total contributions in the 2020 and 2024 election cycles only (so create a dataframe with columns for contributor_name, total_2020, and total_2024). Then calculate a new column showing the change in contribution amount from 2020 to 2024. Which contributors increased their giving the most? Which decreased the most?

A4. 

```{r}
clergy2024change <- clergy_wide |>
  select(-`2016`, -`2018`, -`2022`)

clergy2024change <- clergy2024change|>
  mutate(dollar_change = `2024` - `2020`, pct_change = (`2024` - `2020`) / `2020`)|>
  arrange(desc(dollar_change))

```
#In terms of dollar amount, Doris Cowan, Michael Freeman, and Frank Trotter donated the most. In terms of percentage, David Sword increased the most by 18900% from 50 to 9500. The people that had the biggest drop in dollar amount were John Price, Eldridge Spearmen and Roy Oswald. 


Q5. Now let's work with the United Methodist data. Join the `umdata_people` and `umdata_people_all` datasets using `GCFAId` as the key. Before joining, select only the unique columns from `umdata_people_all` (ConferenceId and ConferenceName) to avoid duplicate columns. After joining, filter for only clergy (not laity) and count them by conference and AccountStatus. Which conferences have the most clergy, and how do Active vs. Deceased clergy compare?

A5.
```{r}
umdata_people <- read_csv('data/umdata_people.csv')
umdata_people_all <- read_csv('data/umdata_people_all.csv')

```

```{r}
people_matched <- umdata_people |>
  inner_join(
    umdata_people_all |> select(GCFAId, ConferenceId, ConferenceName),
    join_by(GCFAId)
  )
clergy <- people_matched |>
filter(str_detect(ClergyOrLay, "Clergy"))

clergy |>
group_by(Conferences)|>
  summarise(count = n())|>
  arrange(desc(count))
clergy|>
  group_by(AccountStatus)|>
  summarise(count = n())|>
  arrange(desc(count))

```
#The largest conference is North Georgia, coming in 300 ahead of the second most, Western North Carolina. Then West Ohio and. Virginia follow just under 2000 clergy. According to the data, there are 18932 active clergy and 9020 deceased clergy on the list. 


Q6. You want to explore whether there's any overlap between clergy who made political contributions and clergy in the United Methodist database. Using the combined contribution data and the UM data, identify any contributors whose names (first and last name) match clergy in the `umdata_people` file. How many matches do you find? What challenges does this kind of name-based matching present? Are there ways to improve the accuracy of matching?

A6. 
```{r}
umdata_people <- umdata_people|>
  mutate(LastName = str_to_upper(LastName), FirstName = str_to_upper(FirstName))
clergy_contribs_combined <- clergy_contribs_combined |>
  rename(FirstName = contributor_first_name, LastName = contributor_last_name)
```
```{r}
umdata_people_lf <-umdata_people |>
   mutate(contributor_name = paste0(LastName, ", ", FirstName))
```
#For the code combining names, I used Claude and this is what I used as my prompt: 
I am using the tidyverse in r. How do I combine a first name column and last name column so that they make a column that is formatted "lastname, first name"

```{r}
clergy_match <- clergy_wide |>
  inner_join(
    umdata_people_lf |> select(contributor_name),
    join_by(contributor_name)
  )
```
#I think I was able to find 518 matches from both datasets. THis may not be a perfect list because there could be people with the same name that appear to be matches, but aren't really. A better way to do this was a shared, unique ID between the datasets which would eliminate this problem. I also see that I have repeat names in my data.  


Q7. Choose your own adventure: Come up with a question that requires you to use at least one join and one aggregation (sum, mean, count, etc.) using either the clergy contribution data, the UM data, or both. Write the question, then write the code to answer it, and provide an answer.

A7: Who is contributing more money, people registered under umdata or people who aren't?

```{r}
nonUMclergy <-clergy_wide |>
  anti_join(umdata_people_lf, join_by(contributor_name))
```
```{r}
clergy_match <- clergy_match|>
  mutate(total = sum(`2016`,`2018`,`2020`,`2022`,`2024`))
nonUMclergy <- nonUMclergy |>
  mutate(total = sum(`2016`,`2018`,`2020`,`2022`,`2024`))
```

```{r}
nonUMclergy|>
  ungroup()|>
summarise(
    avg_cont = mean(total),
    min_cont = min(total),
    max_cont = max(total)
  )
clergy_match|>
  ungroup()|>
  summarise(
     avg_cont = mean(total),
    min_cont = min(total),
    max_cont = max(total)
  )
```
# When I tried to run the code originally here, I ran it without "ungroup" this gave me the mean, min and max for all names. I asked Claude to help me fix by saying "I am using the tidyvrse in r, when I summarise my dataset for mean, min and max, it gives me data for all names, I want one row," and gave it my original code, it suggested to use "ungroup" and that worked. 

```{r}
set.seed(1)
UMsample50 = clergy_match |> sample_n(50)

UMsample50|>
  summarise(
    avg_cont = mean(total),
    min_cont = min(total),
    max_cont = max(total)
  )
NONUMsample50 = nonUMclergy |> sample_n(50)

NONUMsample50 |>
  summarise(
    avg_cont = mean(total),
    min_cont = min(total),
    max_cont = max(total)
  )
```

#I think I am stuck here but I also think I was headed in a good direction. I wanted to see if Methodist clergy contributed more money per person on average than non-Methodist clergy. To do this I first took the whole data of each graph and found the mean, max and min. This gave me obviously inconclusive evidence. Then I wanted to take a random sample of 50 people from each data and then compare the mean, max and min for each sample, which might make it more fair. However, I ran into the issue that I have duplicates which prevents the program from taking a real sample. I am not sure how to combine the names and the dollar amounts for each duplicate, but if I did I think I would have good data. 
